---
title: "Comparing ResMem and CLIP Challengers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comparing ResMem and CLIP Challengers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
params:
  family: red
  preset: homage
css: albers.css
resource_files:
- albers.css
- albers.js
includes:
  in_header: |-
    <script src="albers.js"></script>
    <script>document.addEventListener('DOMContentLoaded',function(){document.body.classList.add('palette-red');});</script>
---

```{r setup, include = FALSE}
if (requireNamespace("ggplot2", quietly = TRUE) && requireNamespace("albersdown", quietly = TRUE)) ggplot2::theme_set(albersdown::theme_albers(family = params$family, preset = params$preset))
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

```{r}
library(memscore)
```

## Why use the benchmarking layer?

If you want more than a single memorability score, `memscore` gives you two
comparison workflows from R. You can benchmark a custom train/val/test manifest
with `memscore_benchmark_manifest()`, or you can run a repeated FIGRIM study
with `memscore_study_figrim()`.

This vignette uses a mocked backend so the examples stay runnable under
`R CMD check`. The return values, file outputs, and summary structures match the
real package interface whether you run through `reticulate` or the CLI bridge.

```{r benchmark-mock-backend, include = FALSE}
options(memscore.cli = "/tmp/memscore")

assignInNamespace(
  ".system2",
  function(command, args, stdout = TRUE, stderr = TRUE) {
    subcommand <- args[[1]]

    if (identical(subcommand, "benchmark-manifest")) {
      json_path <- args[match("--output-json", args) + 1L]
      pred_path <- args[match("--output-predictions", args) + 1L]
      writeLines(
        c(
          "{",
          '  "dataset": "manifest",',
          '  "sizes": {"train": 4, "val": 1, "test": 3},',
          '  "resmem": {"test_metrics": {"spearman": 0.47, "mse": 0.024}},',
          '  "best_clip": {',
          '    "clip_model": "ViT-B-32",',
          '    "regressor": "ridge",',
          '    "test_metrics": {"spearman": 0.66, "mse": 0.013}',
          "  },",
          '  "delta_vs_resmem": {"spearman": 0.19, "mse": -0.011}',
          "}"
        ),
        json_path
      )
      writeLines(
        c(
          "id,image_path,ground_truth,resmem_prediction,clip_prediction",
          "beach,/tmp/beach.jpg,0.72,0.58,0.71",
          "street,/tmp/street.jpg,0.63,0.55,0.65",
          "forest,/tmp/forest.jpg,0.49,0.57,0.50"
        ),
        pred_path
      )
      return(c(
        sprintf("Wrote summary to %s", json_path),
        sprintf("Wrote test predictions to %s", pred_path)
      ))
    }

    if (identical(subcommand, "study-figrim")) {
      json_path <- args[match("--output-json", args) + 1L]
      csv_path <- args[match("--output-csv", args) + 1L]
      writeLines(
        c(
          "{",
          '  "summary_rows": [',
          '    {"protocol": "across_category", "feature_source": "ViT-B-32+RN50+ViT-B-16", "method": "mean_ensemble", "model_mean_spearman": 0.74, "resmem_mean_spearman": 0.45, "mean_delta_spearman": 0.29},',
          '    {"protocol": "within_category", "feature_source": "ViT-B-32+RN50+ViT-B-16", "method": "mean_ensemble", "model_mean_spearman": 0.64, "resmem_mean_spearman": 0.24, "mean_delta_spearman": 0.40}',
          "  ]",
          "}"
        ),
        json_path
      )
      writeLines(
        c(
          "protocol,feature_source,method,model_mean_spearman,resmem_mean_spearman,mean_delta_spearman",
          "across_category,ViT-B-32+RN50+ViT-B-16,mean_ensemble,0.74,0.45,0.29",
          "within_category,ViT-B-32+RN50+ViT-B-16,mean_ensemble,0.64,0.24,0.40"
        ),
        csv_path
      )
      return(c(
        sprintf("Wrote study JSON to %s", json_path),
        sprintf("Wrote study CSV to %s", csv_path)
      ))
    }

    stop("Unexpected mocked subcommand: ", subcommand, call. = FALSE)
  },
  ns = "memscore"
)
```

```{r manifest-data, include = FALSE}
manifest_df <- data.frame(
  path = c(
    "images/beach.jpg",
    "images/museum.jpg",
    "images/street.jpg",
    "images/forest.jpg",
    "images/station.jpg",
    "images/cafe.jpg",
    "images/desert.jpg",
    "images/kitchen.jpg"
  ),
  score = c(0.72, 0.68, 0.63, 0.49, 0.61, 0.58, 0.47, 0.52),
  split = c("train", "train", "train", "train", "val", "test", "test", "test"),
  id = c("beach", "museum", "street", "forest", "station", "cafe", "desert", "kitchen"),
  stringsAsFactors = FALSE
)
manifest_path <- tempfile(fileext = ".csv")
write.csv(manifest_df, manifest_path, row.names = FALSE)
```

## What does a custom benchmark look like?

The shortest useful path is to point `memscore_benchmark_manifest()` at a CSV
with image paths, targets, and split labels.

```{r quick-win}
benchmark <- memscore_benchmark_manifest(
  manifest = manifest_path,
  root = tempdir(),
  clip_models = c("ViT-B-32", "RN50"),
  regressors = "ridge"
)

benchmark$summary$delta_vs_resmem
stopifnot(benchmark$summary$delta_vs_resmem$spearman > 0)
stopifnot(benchmark$summary$delta_vs_resmem$mse < 0)
```

That quick check answers the main question: did the best CLIP challenger beat
the released `ResMem` baseline on the held-out test set?

## What does the manifest need to contain?

The manifest is just a small table with `path`, `score`, and `split`, plus an
optional `id`. Here is the example used above.

```{r show-manifest}
manifest_df
stopifnot(all(c("path", "score", "split", "id") %in% names(manifest_df)))
stopifnot(all(manifest_df$score >= 0 & manifest_df$score <= 1))
stopifnot(all(sort(unique(manifest_df$split)) == c("test", "train", "val")))
```

## How do you inspect per-image outputs?

The benchmark call writes a test-set prediction table that is useful for error
analysis and plotting.

```{r read-predictions}
predictions <- read.csv(benchmark$output_predictions, stringsAsFactors = FALSE)
predictions
stopifnot(file.exists(benchmark$output_predictions))
stopifnot(all(is.finite(predictions$clip_prediction)))
stopifnot(nrow(predictions) == 3)
```

```{r compare-errors}
predictions$resmem_abs_error <- abs(predictions$ground_truth - predictions$resmem_prediction)
predictions$clip_abs_error <- abs(predictions$ground_truth - predictions$clip_prediction)

predictions[, c("id", "resmem_abs_error", "clip_abs_error")]
stopifnot(mean(predictions$clip_abs_error) < mean(predictions$resmem_abs_error))
```

```{r plot-errors, echo = FALSE}
error_means <- c(
  ResMem = mean(predictions$resmem_abs_error),
  CLIP = mean(predictions$clip_abs_error)
)
barplot(
  error_means,
  ylim = c(0, max(error_means) * 1.25),
  col = c("#B8B8B8", "#3F7CAC"),
  ylab = "Mean absolute error",
  main = "Held-out error on the mocked test set"
)
```

The challenger has lower average error in this example, which matches the
summary metrics returned by `benchmark$summary$delta_vs_resmem`.

## How do you summarize a FIGRIM study?

Once you want something closer to the repeated-study results used in this repo,
`memscore_study_figrim()` returns a summary table across protocols and model
configurations.

```{r run-figrim}
study <- memscore_study_figrim(
  figrim_root = tempdir(),
  clip_models = c("ViT-B-32", "RN50", "ViT-B-16"),
  clip_tta = "fivecrop"
)

study_rows <- study$study$summary_rows
study_rows
stopifnot(all(is.finite(study_rows$model_mean_spearman)))
stopifnot(all(study_rows$mean_delta_spearman > 0))
```

```{r plot-figrim, echo = FALSE}
barplot(
  study_rows$mean_delta_spearman,
  names.arg = study_rows$protocol,
  col = c("#3F7CAC", "#2A9D8F"),
  ylab = "Mean delta Spearman vs ResMem",
  main = "Mocked FIGRIM study improvement"
)
abline(h = 0, lty = 2)
```

In a real study, that summary table is the object you would write to disk,
filter, and turn into a paper table or figure.

## What should you reach for next?

- Use `memscore_benchmark_manifest()` when you already have your own split file.
- Use `memscore_benchmark_lamem()` when you want the official LaMem split
  workflow.
- Use `memscore_study_figrim()` when you want repeated FIGRIM comparisons and
  summary tables rather than a single held-out test split.
