---
title: "Getting Started with memscore"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with memscore}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

```{r}
library(memscore)
```

## What does memscore do?

`memscore` gives R users access to the Python `memscore` command line tools
without `reticulate`. You configure a backend once, then call ordinary R
functions such as `memscore_predict()` or `memscore_benchmark_manifest()`.

This vignette uses a mocked backend so the examples run during package checks.
The public R API and return values are the same as in a real session.

```{r mock-backend, include = FALSE}
assignInNamespace(
  ".system2",
  function(command, args, stdout = TRUE, stderr = TRUE) {
    subcommand <- args[[1]]

    if (identical(subcommand, "predict")) {
      return(c(
        "Wrote predictions to /tmp/preds.csv",
        '[{"id":"beach","image_path":"/tmp/beach.jpg","memorability":0.72},',
        '{"id":"forest","image_path":"/tmp/forest.jpg","memorability":0.54}]'
      ))
    }

    if (identical(subcommand, "benchmark-manifest")) {
      json_path <- args[match("--output-json", args) + 1L]
      pred_path <- args[match("--output-predictions", args) + 1L]
      writeLines(
        c(
          "{",
          '  "dataset": "manifest",',
          '  "delta_vs_resmem": {',
          '    "spearman": 0.18,',
          '    "mse": -0.012',
          "  },",
          '  "best_clip": {"clip_model": "ViT-B-32", "regressor": "ridge"}',
          "}"
        ),
        json_path
      )
      writeLines(
        "id,image_path,ground_truth,resmem_prediction,clip_prediction\nbeach,/tmp/beach.jpg,0.70,0.61,0.74",
        pred_path
      )
      return(c(
        sprintf("Wrote summary to %s", json_path),
        sprintf("Wrote test predictions to %s", pred_path)
      ))
    }

    if (identical(subcommand, "study-figrim")) {
      json_path <- args[match("--output-json", args) + 1L]
      csv_path <- args[match("--output-csv", args) + 1L]
      writeLines(
        c(
          "{",
          '  "summary_rows": [',
          '    {"protocol": "across_category", "model_mean_spearman": 0.74},',
          '    {"protocol": "within_category", "model_mean_spearman": 0.64}',
          "  ]",
          "}"
        ),
        json_path
      )
      writeLines(
        "protocol,model_mean_spearman\nacross_category,0.74\nwithin_category,0.64",
        csv_path
      )
      return(c(
        sprintf("Wrote study JSON to %s", json_path),
        sprintf("Wrote study CSV to %s", csv_path)
      ))
    }

    stop("Unexpected mocked subcommand: ", subcommand, call. = FALSE)
  },
  ns = "memscore"
)
```

## How do you configure the backend?

Point R at either a `memscore` executable or a Python interpreter that can run
`python -m memscore`.

```{r configure}
memscore_configure(
  python = "/usr/bin/python3",
  workdir = tempdir()
)

info <- memscore_cli_info()
info
stopifnot(info$source %in% c("path-cli", "python-module"))
```

## How do you score images?

The first workflow is simple scoring. The return value is a data frame with one
row per input image.

```{r predict}
scores <- memscore_predict(c("beach.jpg", "forest.jpg"))
scores
stopifnot(all(c("id", "image_path", "memorability") %in% names(scores)))
stopifnot(all(is.finite(scores$memorability)))
stopifnot(all(scores$memorability >= 0 & scores$memorability <= 1))
```

## How do you benchmark a custom manifest?

The manifest workflow compares the released `ResMem` checkpoint against frozen
CLIP challengers on your own train/val/test split file.

```{r manifest-file, include = FALSE}
manifest_path <- tempfile(fileext = ".csv")
writeLines(
  c(
    "path,score,split,id",
    "beach.jpg,0.70,train,beach",
    "forest.jpg,0.54,val,forest",
    "street.jpg,0.63,test,street"
  ),
  manifest_path
)
```

```{r benchmark-manifest}
benchmark <- memscore_benchmark_manifest(
  manifest = manifest_path,
  clip_models = c("ViT-B-32", "RN50"),
  regressors = "ridge"
)

benchmark$summary$delta_vs_resmem
stopifnot(benchmark$summary$delta_vs_resmem$spearman > 0)
file.exists(benchmark$output_json)
file.exists(benchmark$output_predictions)
```

## How do you launch the FIGRIM study?

For repeated FIGRIM comparisons, `memscore_study_figrim()` exposes the same
workflow that the Python CLI uses.

```{r study-figrim}
study <- memscore_study_figrim(
  figrim_root = tempdir(),
  clip_models = c("ViT-B-32", "RN50", "ViT-B-16"),
  clip_tta = "fivecrop"
)

study$study$summary_rows
stopifnot(nrow(study$study$summary_rows) >= 1)
```

## Where next?

- Use `?memscore_predict` for the scoring interface.
- Use `?memscore_benchmark_manifest` and `?memscore_study_figrim` for the two
  main evaluation workflows.
- See the top-level `README.md` and `BENCHMARKS.md` files in the repository for
  the matching Python commands.
