---
title: "Getting Started with memscore"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with memscore}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
params:
  family: red
  preset: homage
css: albers.css
resource_files:
- albers.css
- albers.js
includes:
  in_header: |-
    <script src="albers.js"></script>
    <script>document.addEventListener('DOMContentLoaded',function(){document.body.classList.add('palette-red');});</script>
---

```{r setup, include = FALSE}
if (requireNamespace("ggplot2", quietly = TRUE) && requireNamespace("albersdown", quietly = TRUE)) ggplot2::theme_set(albersdown::theme_albers(family = params$family, preset = params$preset))
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

```{r}
library(memscore)
```

## What does memscore do?

`memscore` gives R users access to the Python `memscore` package from ordinary
R functions. You can run it in-process through `reticulate`, or fall back to a
CLI bridge when that is more convenient for deployment.

This vignette uses a mocked backend so the examples run during package checks.
The public R API and return values are the same as in a real session.

```{r mock-backend, include = FALSE}
assignInNamespace(
  ".predict_via_reticulate",
  function(...) {
    data.frame(
      id = c("beach", "forest"),
      image_path = c("/tmp/beach.jpg", "/tmp/forest.jpg"),
      memorability = c(0.72, 0.54),
      stringsAsFactors = FALSE
    )
  },
  ns = "memscore"
)

assignInNamespace(
  ".benchmark_manifest_via_reticulate",
  function(...) {
    json_path <- tempfile("memscore-benchmark-", fileext = ".json")
    pred_path <- tempfile("memscore-predictions-", fileext = ".csv")
    writeLines(
      c(
        "{",
        '  "dataset": "manifest",',
        '  "delta_vs_resmem": {',
        '    "spearman": 0.18,',
        '    "mse": -0.012',
        "  },",
        '  "best_clip": {"clip_model": "ViT-B-32", "regressor": "ridge"}',
        "}"
      ),
      json_path
    )
    writeLines(
      "id,image_path,ground_truth,resmem_prediction,clip_prediction\nbeach,/tmp/beach.jpg,0.70,0.61,0.74",
      pred_path
    )
    list(
      summary = jsonlite::read_json(json_path, simplifyVector = TRUE),
      output_json = json_path,
      output_predictions = pred_path,
      stdout = character()
    )
  },
  ns = "memscore"
)

assignInNamespace(
  ".study_figrim_via_reticulate",
  function(...) {
    study_path <- tempfile("memscore-figrim-", fileext = ".json")
    csv_path <- tempfile("memscore-figrim-", fileext = ".csv")
    writeLines(
      c(
        "{",
        '  "summary_rows": [',
        '    {"protocol": "across_category", "model_mean_spearman": 0.74},',
        '    {"protocol": "within_category", "model_mean_spearman": 0.64}',
        "  ]",
        "}"
      ),
      study_path
    )
    writeLines(
      "protocol,model_mean_spearman\nacross_category,0.74\nwithin_category,0.64",
      csv_path
    )
    list(
      study = jsonlite::read_json(study_path, simplifyVector = TRUE),
      output_json = study_path,
      output_csv = csv_path,
      stdout = character()
    )
  },
  ns = "memscore"
)
```

## How do you configure the backend?

The recommended path is to point R at a Python interpreter and set
`backend = "reticulate"`.

```{r configure}
memscore_configure(
  python = "/usr/bin/python3",
  backend = "reticulate",
  workdir = tempdir()
)

getOption("memscore.backend")
stopifnot(identical(getOption("memscore.backend"), "reticulate"))
```

If you need a pure command-line fallback instead, you can force that explicitly.

```{r configure-cli, eval = FALSE}
memscore_configure(
  cli = "/path/to/memscore",
  backend = "cli",
  workdir = "/path/to/memscore"
)
```

## How do you score images?

The first workflow is simple scoring. The return value is a data frame with one
row per input image.

```{r predict}
scores <- memscore_predict(c("beach.jpg", "forest.jpg"))
scores
stopifnot(all(c("id", "image_path", "memorability") %in% names(scores)))
stopifnot(all(is.finite(scores$memorability)))
stopifnot(all(scores$memorability >= 0 & scores$memorability <= 1))
```

## How do you benchmark a custom manifest?

The manifest workflow compares the released `ResMem` checkpoint against frozen
CLIP challengers on your own train/val/test split file.

```{r manifest-file, include = FALSE}
manifest_path <- tempfile(fileext = ".csv")
writeLines(
  c(
    "path,score,split,id",
    "beach.jpg,0.70,train,beach",
    "forest.jpg,0.54,val,forest",
    "street.jpg,0.63,test,street"
  ),
  manifest_path
)
```

```{r benchmark-manifest}
benchmark <- memscore_benchmark_manifest(
  manifest = manifest_path,
  clip_models = c("ViT-B-32", "RN50"),
  regressors = "ridge"
)

benchmark$summary$delta_vs_resmem
stopifnot(benchmark$summary$delta_vs_resmem$spearman > 0)
file.exists(benchmark$output_json)
file.exists(benchmark$output_predictions)
```

## How do you launch the FIGRIM study?

For repeated FIGRIM comparisons, `memscore_study_figrim()` exposes the same
workflow regardless of whether you route through `reticulate` or the CLI.

```{r study-figrim}
study <- memscore_study_figrim(
  figrim_root = tempdir(),
  clip_models = c("ViT-B-32", "RN50", "ViT-B-16"),
  clip_tta = "fivecrop"
)

study$study$summary_rows
stopifnot(nrow(study$study$summary_rows) >= 1)
```

## Where next?

- Use `?memscore_predict` for the scoring interface.
- Use `?memscore_benchmark_manifest` and `?memscore_study_figrim` for the two
  main evaluation workflows.
- See the top-level `README.md` and `BENCHMARKS.md` files in the repository for
  the matching Python commands.
